---
title: "NYPD Shooting Incidents - Analysis"
author: "Praveen Myakala"
date: "2023-10-23"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}

library(tidyverse)
```

## Get NYPD Shooting Incidents & Display Summary
#### NYPD Shooting Incident Data (Historic) from the website: https://catalog.data.gov/datase

```{r get_data, echo=TRUE}
# Download data
nypd_shooting_incidents <- read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv")
```

## Displaying Summary
```{r summary, echo=TRUE}
# Display summary
summary(nypd_shooting_incidents)
```

## Preparation and Cleaning
#### Mainly removing duplicates by INCIDENT_KEY and selecting only required columns to analysis with mutate
```{r prepare, echo=TRUE}
# Distinct data with Mutate
nypd_shooting_incidents <- nypd_shooting_incidents %>%
  distinct(INCIDENT_KEY, .keep_all = TRUE) %>%
  select(INCIDENT_KEY, PRECINCT, OCCUR_DATE, OCCUR_TIME, BORO, PERP_AGE_GROUP, PERP_SEX, PERP_RACE, VIC_AGE_GROUP, VIC_RACE, VIC_SEX,) %>%
  mutate(OCCUR_DATE = mdy(OCCUR_DATE)) %>%
  mutate(YEAR = year(OCCUR_DATE))
```

## Summary after cleaning
```{r summary_clean_data, echo=TRUE}
# Displaying summary
summary(nypd_shooting_incidents)
```


#### Group by BORO and calculate the sum of incidents

```{r by_boro, echo=TRUE}
# Grouping incidents by BORO
shootings_incidents_by_boro <- nypd_shooting_incidents %>% 
  group_by(BORO) %>%  
  summarize(incidents = n())

```

## Displaying Summary after grouping incidents by BORO
```{r summary_of_boro, echo=TRUE}
# Summary
summary(shootings_incidents_by_boro)
```



```{r by_boro_graph, echo=TRUE}

# ggplot code for bar graph

ggplot(shootings_incidents_by_boro, aes(x=BORO, y=incidents, fill=BORO)) +
  geom_bar(stat="identity") + 
  xlab("BOROUGH") + ylab("INCIDENCES") +
  ggtitle("Shooting Incidents By BOROUGHS") +
  theme_minimal()

```


## NYPD Shooting incidents by victim gender

```{r Vicitims_by_gender, echo=TRUE}
# Graph with victim gender

nypd_shooting_incidents %>% 
  filter(VIC_SEX %in% c("F", "M")) %>%
  count(YEAR, VIC_SEX) %>%
  ggplot(aes(x = YEAR, y = n, color = VIC_SEX)) + 
  geom_line() + 
  labs(x = "Year", y = "Shooting Incidents", title = "") + 
  theme_minimal() + 
  scale_color_manual(values = c("M" = "blue", "F" = "pink")) + 
  theme_minimal()
```

## NYPD Shooting incidents over the course of the day by precint

```{r over_the_day, echo=TRUE}

# Graph incidents over time period

ggplot(nypd_shooting_incidents, aes(x = OCCUR_TIME, y = PRECINCT, color = BORO)) + 
  geom_point(size = 0.2) + 
  theme_minimal()
```
